This is the story of unexpected results, obtained by an amateur
student with a deep curiosity for mathematics.
Although, not a specialist in mathematics,
our young explorer always possessed a strong sense of mathematical beauty and aesthetics.
The mathematical knowledge of the individual was limited by undergraduate level course, which includes the basics of
matrix operations, basic calculus, and elementary linear algebra.
One day, our student found himself observing the tables of finite differences,
in particular finite differences of cubes

\input{sections/figures/01_fig_finite_differences_cubes}
The first question that visited curious mind was
\begin{problem}
    \label{quest:interpolation-cubes}
    How to reconstruct the value of $n^3$ from the values of its finite differences?
\end{problem}
Precisely, the inquiry was to find a way to reconstruct the values of the sequence $\{0, 1, 8, 27, 64, \ldots\}$
given the values of its finite differences.
Clearly, the inquiry refers to the process of interpolation.

In its essence, the problem of interpolation is so old that it can be traced back
to ancient Babylonian and Greek times, several centuries BC and first centuries AD~\cite{gautschi2012interpolation}.
Interpolation -- is the process of finding new data points based on the range of a discrete set
of known data points.
Interpolation, as we know it today, has been developed in between 1674--1684 by Sir Isaac Newton.
His works in interpolation are frequently referenced as
the foundation of classical interpolation theory~\cite{meijering2002chronology}.
For instance, Newton's interpolation formula addresses the problem~\eqref{quest:interpolation-cubes} immediately,
because
\begin{align*}
    27 = 6 \binom{3}{3} + 6 \binom{3}{2} + 1 \binom{3}{1} + 0\binom{3}{0} = 6 + 18 + 3
\end{align*}
where $6,6,1,0$ corresponds to the first row in table~\eqref{tab:finite-differences-cubes}.
In general,
\begin{align*}
    n^3 = 6 \binom{n}{3} + 6 \binom{n}{2} + 1\binom{n}{1} + 0 \binom{n}{0}
\end{align*}
because $f(x) = \sum_{k=0}^{d} \Delta^{d-k} f(0) \binom{x}{d-k}$, see~\cite[~p. 190]{graham1994concrete}.

Great!
But there is one thing, the student which stated the problem~\eqref{quest:interpolation-cubes}
had no clue about interpolation theory at all.
What was the action taken next?
Exactly, our inquirer decided to re-invent the solution himself, being driven by the pure feeling of mystery.
Student's mind was occupied by only a single thought:
\textit{All mathematical truths exist timelessly, we only reveal and describe them.}
That mindset inspired our researcher to start his own mathematical journey.

By observing the table of finite differences~\eqref{tab:finite-differences-cubes} we can notice that
the first order finite difference of cubes may be expressed in terms of its
third order finite differences $\Delta^3(n^3) = 6$, as follows
\begin{align*}
    \begin{split}
        \Delta(0^3) &= 1+6 \cdot 0 \\
        \Delta(1^3) &= 1+6\cdot0+6\cdot1 \\
        \Delta(2^3) &= 1+6\cdot0+6\cdot1+6\cdot2 \\
        \Delta(3^3) &= 1+6\cdot0+6\cdot1+6\cdot2+6\cdot3 \\
        &\; \; \vdots \\
        \Delta(n^3) &= 1+6\cdot0+6\cdot1+6\cdot2+6\cdot3 + \cdots + 6n
    \end{split}
\end{align*}
By using sigma notation, we get
\begin{align*}
    \Delta(n^3) = 1+6\cdot0+6\cdot1+6\cdot2+6\cdot3+\cdots+6\cdot n = 1 + 6 \sum_{k=0}^{n} k
\end{align*}

However, there is a more beautiful way to prove that $\Delta(n^3) = 1 + 6 \sum_{k=0}^{n} k$.
We refer to one of the finest articles in the area of polynomials and power sums,
that is \textit{Johann Faulhaber and sums of powers} written by Donald Knuth~\cite{knuth1993johann}.
We now focus on the odd power identities shown at~\cite[~p. 9]{knuth1993johann}
\begin{align*}
    n^1 &= \binom{n}{1} \\
    n^3 &= 6 \binom{n+1}{3} + \binom{n}{1} \\
    n^5 &= 120 \binom{n+2}{5} + 30 \binom{n+1}{3} + \binom{n}{1}
\end{align*}

It is quite interesting that the identity in terms of triangular numbers $\binom{n+1}{2}$
and finite differences of cubes becomes more obvious
\begin{align*}
    \Delta n^3
    = (n+1)^3 - n^3
    =  6 \binom{n+1}{2} + \binom{n}{0}
\end{align*}
It easy to see that
\begin{align*}
    \Delta n^3
    = \left[ 6 \binom{n+2}{3} + \binom{n+1}{1} \right] - \left[ 6 \binom{n+1}{3} + \binom{n}{1} \right]
    = 6 \binom{n+1}{2} + \binom{n}{0}
\end{align*}
because $\binom{n}{k} = \binom{n-1}{k} + \binom{n-1}{k-1}$.

Moreover, the concept above allows us to reach $r$-fold power sums $\sum^r k^{2m+1}$
or finite differences $\Delta^r k^{2m+1}$ by altering binomial coefficients indices.
Quite strong and impressive.
\begin{align*}
    \Sigma^{r} k^3 &=  6 \binom{n+1+r}{3+r} + \binom{n+r}{1+r} \\
    \Delta^{r} k^5 &= 120 \binom{n+2-r}{5-r} + 30 \binom{n+1-r}{3-r} + \binom{n-r}{1-r}
\end{align*}

We can observe that triangular numbers $\binom{n+1}{2}$ are equivalent to
\begin{align*}
    \binom{n+1}{2} = \sum_{k=0}^{n} \binom{k}{1} = \sum_{k=0}^{n} k
\end{align*}
because $\binom{n+1}{m+1} = \sum_{k=0}^{n} \binom{k}{m}$.
This leads us to the fact that
\begin{align*}
    \Delta n^3 = (n+1)^3 - n^3 = 1 + 6 \sum_{k=0}^{n} k
\end{align*}
An experienced mathematician would immediately notice a spot to apply Faulhaber's formula~\cite{beardon1996sums}
to get the closed form of the sum $\sum_{k=0}^{n} k$
\begin{align*}
    \sum_{k=0}^{n} k = \frac{1}{2}(n+n^2)
\end{align*}
Thus, the finite difference $\Delta(n^3)$ takes the well-known form,
which matches Binomial theorem~\cite{abramowitz1988handbook}
\begin{align*}
    \Delta(n^3)
    = 1 + 6 \left[ \frac{1}{2}(n+n^2) \right]
    = 1 + 3 n + 3 n^2
    = \sum_{k=0}^{2} \binom{3}{k} n^k
\end{align*}
And\ldots that could be the end of the story, isn't it?
Because all what remains is to say that
\begin{align*}
    n^3
    = \sum_{k=0}^{n-1} (k+1)^3 - k^3
    = \sum_{k=0}^{n-1} \left( 1 + 6 \sum_{t=0}^{k} t \right)
    = \sum_{k=0}^{n-1} 1 + 3 k + 3 k^2
\end{align*}
Thus, the formula for cubes $n^3$ was derived, and thus,
our protégée's problem~\eqref{quest:interpolation-cubes} has been solved successfully.
Because we have found a formula that recovers the cubes $n^3$ from their finite differences.
Indeed, $5^3 = \sum_{k=0}^{4} 1 + 3 k + 3 k^2 = 1 + 7 + 19 + 37 + 61 = 125$ which is exactly
the sum of terms of the first column of~\eqref{tab:finite-differences-cubes}.

However, not this time.
Luckily enough (say), the student who stated the problem~\eqref{quest:interpolation-cubes}
wasn't really aware of the approaches above neither.
What a lazy student!
Probably, that's exactly the case when unawareness leads to a fresh sight into classical problems,
leading to unexpected results and new insights.
Instead, our investigator spotted a little bit different pattern in
$\Delta n^3= 6 \binom{n+1}{2} + \binom{n}{0} = 1 +  6 \sum_{k=0}^{n} k$.

Consider the polynomial $n^3$ as sum of its finite differences
$\Delta n^3 = 1 +  6 \sum_{k=0}^{n} k$
\begin{align*}
    n^3
    &= [1+6\cdot0] \\
    &+ [1+6\cdot0+6\cdot1] \\
    &+ [1+6\cdot0+6\cdot1+6\cdot2] + \cdots \\
    &+ [1+6\cdot0+6\cdot1+6\cdot2+\cdots+6\cdot(n-1)]
\end{align*}
We can observe that the term $1$ appears $n$ times, while the term $6\cdot0$ appears $n-0$ times.
The term $6\cdot1$ appears $n-1$ times and so on.
By combining the recurring common terms, we get
\begin{align*}
    n^3 = n
    &+ [(n-0) \cdot 6 \cdot 0] \\
    &+ [(n-1)\cdot6\cdot1] \\
    &+ [(n-2)\cdot6\cdot2] + \cdots \\
    &+ [(n-k)\cdot 6 \cdot k] + \cdots \\
    &+ [1\cdot6\cdot(n-1)]
\end{align*}
By applying compact sigma sum notation yields an identity for cubes $n^3$
\begin{align*}
    n^3 = n + \sum_{k=0}^{n-1} 6k(n-k)
\end{align*}
We can freely move the term $n$ under the summation, because there are exactly $n$ iterations.
Therefore,
\begin{align*}
    n^3 = \sum_{k=0}^{n-1} 6k(n-k) + 1
\end{align*}
By inspecting the expression $6k(n-k) + 1$, we can notice that it is symmetric over $k$.
Let be $T_{1} (n,k) = 6k(n-k) + 1$ then
\begin{align*}
    T_{1} (n,k) = T_{1} (n,n-k)
\end{align*}
This symmetry allows us to alter summation bounds easily.
Hence,
\begin{align*}
    n^3 = \sum_{k=1}^{n} 6k(n-k) + 1
\end{align*}
By arranging the values of $T_{1} (n,k)$ as tabular, we see that the identities above are indeed true.
\input{sections/figures/02_fig_triangle_row_sums_give_cubes}
The following recurrence holds for $T_{1} (n,k)$
\begin{align*}
    T_{1} (n, k) = 2T_{1} (n-1, k) - T_{1} (n-2, k)
\end{align*}
Which is indeed true, because
\begin{align*}
    T_{1} (5, 2) = 2 \cdot 25 - 13 = 37
\end{align*}
Finally, our curious learner has reached the first milestone, by finding an
unusual solution to the problem~\eqref{quest:interpolation-cubes}.
What an excitement it was!
However, the excitement would not take too long.
Indeed, curiosity is not something that can be fulfilled completely,
and thus new questions arise.
Somehow, the inquirer got a strong feeling that more general
law hides behind the identity $n^3 = \sum_{k=1}^{n} 6k(n-k) + 1$.
That was quite intuitive.
Fair enough that the next problem was
\begin{problem}
    Is there a generalization of identity $n^3 = \sum_{k=1}^{n} 6k(n-k) + 1$ for higher-degree powers?
    \label{question:higher_powers}
\end{problem}

However, it was not so easy for our young explorer to generalize cube identity for higher powers $n^4$ or $n^5$
by observing the tables of finite differences, as before.
The previous approach to express the difference of cubes $\Delta n^3$ in terms of
$\Delta^3 n^3 = 6$, reaching the identity $n^3 = \sum_k 6k (n-k) +1$ --- was not successful.
Moreover, it wasn't even clear what is the general form of an identity we are looking for.
Thus, the problem~\eqref{question:higher_powers} was shared with the mathematical community,
and there was an answer.
