Once there was a curious student, curious in general and especially in mathematics.
He was not an educated mathematician, but he was strongly captured by the mathematical
beauty and aesthetics.

One day this student was having fun with tables of finite differences, precisely finite differences of cubes.
By observing the table
\input{sections/figures/01_fig_finite_differences_cubes}
The first question that triggered his mind was
\begin{question}
    \label{quest:interpolation-cubes}
    How to reconstruct the value of $n^3$ from its finite differences?
\end{question}
The question is quite old, it can be traced back to ancient Babylonian and Greek times,
several centuries BC and first centuries AD [citation].
The process is called interpolation which is a process
of finding new data points based on the range of a discrete set
of known data points.
Interpolation we know nowadays was developed in 1674--1684 by Issac Newton
in his works referenced as foundation of classical interpolation theory [citation].

Great!
But there is one thing, the student who has risen the question~\eqref{quest:interpolation-cubes}
had no clue about interpolation theory at all.
What he decided then?
Exactly, he decided to try to re-invent interpolation formula himself,
fueled by the purest feeling of mystery.
His mind was occupied by only a single thought:
\textit{
    All mathematical truths exist timelessly, we only reveal and describe them.
}
That mindset inspired our student to start his own mathematical journey.

By observing the table of finite differences~\eqref{tab:finite-differences-cubes} we can notice that
the first order finite difference of cubes may be expressed in terms of its
third order finite difference $\Delta^3(n^3) = 6$, as follows
\begin{align*}
    \begin{split}
        \Delta(0^3) &= 1+6 \cdot 0 \\
        \Delta(1^3) &= 1+6\cdot0+6\cdot1 \\
        \Delta(2^3) &= 1+6\cdot0+6\cdot1+6\cdot2 \\
        \Delta(3^3) &= 1+6\cdot0+6\cdot1+6\cdot2+6\cdot3 \\
        &\; \; \vdots \\
        \Delta(n^3) &= 1+6\cdot0+6\cdot1+6\cdot2+6\cdot3 + \cdots + 6n
    \end{split}
\end{align*}
By using sigma notation for sums we get
\begin{align*}
    \Delta(n^3) = 1+6\cdot0+6\cdot1+6\cdot2+6\cdot3+\cdots+6\cdot n = 1 + 6 \sum_{k=0}^{n} k
\end{align*}

However, there is a more beautiful way to prove that $\Delta(n^3) = 1 + 6 \sum_{k=0}^{n} k$.
We refer to one of the finest articles in the area of polynomials and power sums,
that is \textit{Johann Faulhaber and sums of powers} written by Donald Knuth~\cite{knuth1993johann}.
Indeed, this article is a great source to reach piece of mind in mathematics.
We now focus on the following odd power identities shown on p.9 of~\cite{knuth1993johann}
\begin{align*}
    n^1 &= \binom{n}{1} \\
    n^3 &= 6 \binom{n+1}{3} + \binom{n}{1} \\
    n^5 &= 120 \binom{n+2}{5} + 30 \binom{n+1}{3} + \binom{n}{1}
\end{align*}

It is particularly interesting that well-known identity that connects triangular numbers $\binom{n+1}{2}$
and finite differences of cubes becomes clear and obvious
\begin{align*}
    \Delta n^3 = (n+1)^3 - n^3
    =  6 \binom{n+1}{2} + \binom{n}{0}
\end{align*}
It easy to see that
\begin{align*}
    \Delta n^3
    = \left[ 6 \binom{n+2}{3} + \binom{n+1}{1} \right] - \left[ 6 \binom{n+1}{3} + \binom{n}{1} \right]
    = 6 \binom{n+1}{2} + \binom{n}{0}
\end{align*}
because $\binom{n}{k} = \binom{n-1}{k} + \binom{n-1}{k-1}$.

Moreover, the concept above allows to reach $N$-fold power sums $\sum^N k^{2m+1}$
or finite differences $\Delta^N k^{2m+1}$ of odd powers by simply altering
binomial coefficients indexes.
Quite strong and impressive.

We can observe that triangular numbers $\binom{n+1}{2}$ are equivalent to
\begin{align*}
    \binom{n+1}{2} = \sum_{k=0}^{n} k
\end{align*}
because $\binom{n+1}{m+1} = \sum_{k=0}^{n} \binom{k}{m}$.
This leads to the identity in finite differences of cubes
\begin{align*}
    \Delta n^3 = (n+1)^3 - n^3 = 1 + 6 \sum_{k=0}^{n} k
\end{align*}
An experienced mathematician would immediately notice a spot to apply Faulhaber's formula~\cite{beardon1996sums}
to expand the term $\sum_{k=0}^{n} k$
\begin{align*}
    \sum_{k=0}^{n} k = \frac{1}{2}(n+n^2)
\end{align*}
Thus, the finite difference $\Delta(n^3)$ takes a well-known form such
that matches Binomial theorem~\cite{abramowitz1988handbook}
\begin{align*}
    \Delta(n^3)
    = 1 + 6 \left[ \frac{1}{2}(n+n^2) \right]
    = 1 + 3 n + 3 n^2
    = \sum_{k=0}^{2} \binom{3}{k} n^k
\end{align*}
And\ldots that could be the end of the story, isn't it?
Because all what remains is to say that
\begin{align*}
    n^3
    = \sum_{k=0}^{n-1} (k+1)^3 - k^3
    = \sum_{k=0}^{n-1} \left( 1 + 6 \sum_{t=0}^{k} t \right)
    = \sum_{k=0}^{n-1} 1 + 3 k + 3 k^2
\end{align*}
Thus, the polynomial $n^3$ is interpolated successfully, and thus, the question~\eqref{quest:interpolation-cubes} is answered
positively.
Because we have successfully found the function that matches $n^3$ from the values of its finite differences from the
table~\eqref{tab:finite-differences-cubes}.

However, not this time.
Luckily enough (say), the student who has stated the question~\eqref{quest:interpolation-cubes}
wasn't really aware of the approaches above.
Probably, that's exactly the case when unawareness leads to a fresh sight to century-old questions,
leading to unexpected results and new insights.
%Instead, our student tried something different.
%More precisely, he rewrote the cubes identity $n^3 = \sum_{k=0}^{n-1} \left( 1 + 6 \sum_{t=0}^{k} t \right)$
%explicitly
%\begin{align*}
%    n^3 &= [1+6\cdot0]+[1+6\cdot0+6\cdot1]+[1+6\cdot0+6\cdot1+6\cdot2] \\
%    &+ \cdots + [1+6\cdot0+6\cdot1+6\cdot2+\cdots+6\cdot(n-1)]
%\end{align*}
%\begin{align*}
%    n^3 = n &+ [(n-0) \cdot 6 \cdot 0] + [(n-1)\cdot6\cdot1] + [(n-2)\cdot6\cdot2] \\
%    &+ \cdots + [(n-k)\cdot 6 \cdot k] + \cdots + [1\cdot6\cdot(n-1)] \nonumber
%\end{align*}
%By applying compact sigma sum notation yields an identity for cubes $n^3$
%\begin{align*}
%    n^3 = n + \sum_{k=0}^{n-1} 6k(n-k)
%\end{align*}
%The term $n$ in the sum above can be moved under sigma notation, because there is exactly $n$ iterations, therefore
%\begin{align*}
%    n^3 = \sum_{k=0}^{n-1} 6k(n-k) + 1
%\end{align*}
%By inspecting the expression $6k(n-k) + 1$ we iterate under summation,
%we can notice that it is symmetric over $k$.
%Let be $T(n,k) = 6k(n-k) + 1$ then
%\begin{align*}
%    T(n,k) = T(n,n-k)
%\end{align*}
%This symmetry allows us to alter summation bounds easily, so that
%\begin{align*}
%    n^3 = \sum_{k=1}^{n} 6k(n-k) + 1
%\end{align*}

%However, as was said, I was not the experienced one mathematician back then,
%so that I reviewed the relation~\eqref{eq:general-cube-eq} from a little bit different perspective.
%Not following the convenient solution~\eqref{eq:cubes-difference-binomial-theorem},
%I have introduced the explicit formula for cubes, using~\eqref{eq:cubes_interpolation}
%\begin{align}
%    \label{eq:rearrangement_to_get_cubes}
%    n^3 &= [1+6\cdot0]+[1+6\cdot0+6\cdot1]+[1+6\cdot0+6\cdot1+6\cdot2]+\cdots \nonumber \\
%    &+[1+6\cdot0+6\cdot1+6\cdot2+\cdots+6\cdot(n-1)]
%\end{align}
%Then, rearranging the terms in equation~\eqref{eq:rearrangement_to_get_cubes} so that it turns into summation
%in terms of $k (n-k)$
%\begin{equation*}
%    \begin{split}
%        n^3 &= n + [(n-0) \cdot 6 \cdot 0] + [(n-1)\cdot6\cdot1] + [(n-2)\cdot6\cdot2] + \cdots \\
%        &\cdots + [(n-k)\cdot 6 \cdot k] + \cdots + [1\cdot6\cdot(n-1)]
%    \end{split}
%\end{equation*}
%By applying compact sigma notation and moving $n$ under summation because there is exactly $n$ iteration, yields
%\begin{equation}
%    \label{eq:cube_identity}
%    n^3 = n + \sum_{k=1}^{n} 6k(n-k); \quad \quad n^3 = \sum_{k=1}^{n} 6k(n-k) + 1
%\end{equation}
%\input{sections/figures/02_fig_triangle_row_sums_give_cubes}
%Therefore, we have reached our base case by successfully interpolating the polynomial $n^3$.
%Fairly enough that the next curiosity would be
%\begin{question}
%    Well, if the relation~\eqref{eq:cube_identity} true for the polynomial $n^3$,
%    then is it true that~\eqref{eq:cube_identity} can be generalized for higher powers, e.g.\ for $n^4$ or $n^5$ similarly?
%    \label{question:higher_powers}
%\end{question}
%That was the next question, however without any expectation of the final form of generalized formula.
%Long story short, the answer to this question is also \textit{Yes}, by utilizing certain approaches
%in terms of systems of linear equations or recurrence formula, which is discussed.
%
%Let us begin from the background and history overview of systems of linear equations approach.
